---
title: "Zeyd Khalil HW9, October 22, 2020"
output: pdf_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(nycflights13)
d <- data(package = "nycflights13")
library(Lahman)
library(babynames)
library(nasaweather)
```  

# Exercise 1 - Identify the primary keys in the following datasets (1.5 points: 1/2 pt. for each dataset). Be sure to show that you have the primary key by showing there are no duplicate entries.

* Lahman::Batting
* babynames::babynames
* nasaweather::atmos

```{r}
BattingKey <- Batting %>%
    mutate(PrimaryKey = row_number()) %>%
    select(PrimaryKey, everything())

BattingKey %>% 
  count(PrimaryKey) %>% 
  filter(n>1)
```

> I found out that the Batting table does not have a Primary Key, because all the variables repeat themselves more than once. What I did was I added a surrogate key to Batting using mutate and row_number().

```{r}
babynames %>% 
  count(year, sex, name, n) %>% 
  filter(nn>1)
```

> The Primary Key in babynames is the combination of year, sex, name and n. When there's more than one Primary Key, we usually call this a Composite Key. 

```{r}
atmos %>% 
  count(lat, long, year, month) %>% 
  filter(n>1)
``` 

> The Primary Key in atmos is the Composite Key that has lat, long, year, and month. 

# Exercise 2 - What is the relationship between the Batting, Master, and Salaries tables in the Lahman package? What are the keys for each dataset and how do they relate to each other?

> The relationship between the Batting, Master, and Salaries tables is that they all have a playerID variable. 

```{r}
Master %>% 
  count(playerID) %>% 
  filter(n>1)
Salaries %>% 
  count(playerID) %>% 
  filter(n>1) %>% 
  head()
Batting %>% 
  count(playerID) %>% 
  filter(n>1) %>% 
  head()
```

> The primary key in Master is playerID. However, that is not the case in Salaries and Batting. In fact, Salaries and Batting both do not have a primary key, because all their variables repeat themselves more than once. 

# Exercise 3 - (2 points) Use an appropriate join to add a column containing the airline name to the flights dataset. Be sure to put the carrier code and name in the first two columns of the result so we can see them. Save the result as flights2.  

```{r}
flights2 <- full_join(airlines, flights)

flights2
```  

# Exercise 4 - Use an appropriate join to add the airport name to the flights2 dataset you got in exercise 3. The codes and names of the airports are in the airports dataset of the nycflights13 package. Put the carrier and carrier name first followed by the destination and destination name, then everything else.  

```{r}
flights3 <- flights2 %>% 
  left_join(airports, c("dest" = "faa")) %>% 
  rename(c("dest_name" = "name.y"), c("dest." = "origin"), 
         c("carrier_name" = "name.x"))

flights3 %>% select(carrier, carrier_name, dest., dest_name, everything())
```  

# Exercise 5 - Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays.

* Use the size or colour of the points to display the average delay for each airport.

* Add the location of the origin and destination (i.e. the lat and lon) to flights.

* Compute the average delay by destination.

```{r}
flights3 <- flights3 %>% 
  group_by(dest) %>% 
  summarize(avg_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  select(everything(), avg_delay)  

head(flights3, n=10)

flights4 <- flights3 %>% 
  full_join(airports, c("dest" = "faa"))

flights4 <- flights4 %>% 
  na.omit()

flights4 %>% 
  ggplot(aes(lon, lat, color = avg_delay)) + 
  borders("state") + 
  geom_point() + 
  coord_quickmap()
```   

# Exercise 6 - Use a set operation function to find which airport codes from flights are not in the airports dataset. 

```{r}
flightsOrigin <- flights %>% 
  select(origin)

airportsFaa <- airports %>% 
  select(faa) %>% 
  rename(c("origin" = "faa"))

setdiff(flightsOrigin, airportsFaa)
```

> This shows that there are no airport codes from the flights dataset that are not in the airports dataset

```{r}
setdiff(airportsFaa, flightsOrigin)
```  

> This, however, shows that there are 1445 airport codes in the airports dataset that are not in the flights dataset. 

